MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-12-11 chapters/MetaTracts.tex
minor changes

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute a similar metric. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$). The following measures are defined in~\cite{Frangi1998}.  

@@

The next set of tasks evaluated the spatial context perceived from MetaTracts.\replacedWith{ We achieve this by asking the evaluators}{The participants were asked} to select the fiber bundle which is largest, smallest, or the most similar to a given one from a \removed{sub}set of fiber bundles. Due to the problems of occlusions and similar grayscale values these tests proven to be more difficult to perform successfully with volume rendering of the plain data, compared to MetaTracts. In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
Finally, the majority of participants agreed that the extracted meshes provides better spatial context than volume rendering.

\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}[tb]
\centering
	\includegraphics[width=\linewidth,  trim = 0mm 0mm 0mm 00mm, clip]{images_pvis/figure9_AMA.pdf}
	 \vspace{-1.5em}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Figure~\ref{fig:orientation_clustering}d, clustered into 10 clusters. The unit for length is the grid cube length.  Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10. }
	\label{fig:len_dist_crop16}
	\vskip-0.2cm
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$. Large $h$ keeps the major bundles intact. For example, in  Figure~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Figure~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown \added{(avoiding over-segmentation)}.
The following parameters were fixed for all the tests. We set the reliable Hessian threshold $R_{H}$ to be 0.3 for all our tests. A $R_{H}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection. \replacedWith{On the other hand, a}{A} very high $ R_{H}$ would lead to a decline in number of MetaTracts produced. The $\alpha$ and $\beta$ in $R_{H}$ are as explained in Frangi et al. \cite{Frangi1998} and set to 0.5. The length and the radius parameters for the cylinders of MetaTracts decide how coarse the fiber cylinders are. These are dependent on the underlying fiber characteristic and the weaving pattern. Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber. We used 10.00 and 2.00 for length and radius (measured in grid voxel size), respectively for all tests.  
Simpler geometry (\replacedWith{example dataset 2}{e.g., D2}) was experimentally found to handle larger cylinders better. $\eta$ in sec.~\ref{subsec:dist_clustering} decides how quickly the hierarchical clustering converges, experimentally 0.3 to 0.6 gave similar results. 
 %Short tracts of less than 0.01 times the smallest axes size are prone to have erroneous local orientation and are therefore removed. 
 $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays; we have used [7-10] and half the length of cylinder (5.0). Our number of fiber bundle directions are limited. Thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well. We compared  $m=3-7$ experimentally without any change in results. 
%\begin{figure}[tb]
%	\centering
%	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
%	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
%	\label{fig:comparison}
%\end{figure}
%

%\begin{figure}
%\centering
% \subfloat[$\eta=7$]{\includegraphics[width=0.3\linewidth]{imagesMetaTracts/Rplot-ndim3k2}}
%  \subfloat[$\eta=3$]{\includegraphics[width=0.3\linewidth]{imagesMetaTracts/Rplot-ndim7k2}}
%  \caption{Embedding from $\mathbb{R}^m\to \mathbb{R}^\eta$, here m is the number of MetaTracts, 580 for this example. Followed by clustering into K=2 clusters.}
%  \label{fig:eta}
%\end{figure}
%\begin{figure}[htp]
%\centering
% \includegraphics[width=\linewidth]{imagesMetaTracts/kul-results.eps}
%  \caption{Results from orientation based dimensionality reduction and clustering. A is the original data set, B shows the results of fiber tracking. C shows the results of orientation based clustering and D shows the results of the proximity based clustering. }\label{fig:kull-full}
%\end{figure}
\section{Limitations and future work}
\label{sec:conclusions}
\added{
 A key assumption of the method is ``connectivity'' (Sec.~\ref{sec:char_data}).
If the ``connectivity'' is not satisfied due to noise in the image, the then MetaTracts will be inaccurate. 
The clustering process also assumes that the fiber bundles have a minimum width.
}
\removed{
In this paper, we introduce a framework to extract and visualize fiber bundles in composite materials. We show that our framework works at comparatively low resolution and with dense fiber arrangements (when extracting single fibers might not be possible). It handles complex fiber patterns such as ``cross overs'' and ``braiding''. We perform an user evaluation to indicate the usefulness of the method.
}
In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up extraction times.
%
%\begin{figure}
%  \centering
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-scalar.eps}}
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-tracks-1.eps}}
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-tracks-clus-a.eps}}
%    \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-tracks-clus-b.eps}}
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-labels-all.eps}}
%  \caption{Dense data set with bundles which are flat and thin. (a) shows the results original dataset, (b) shows the results of metaTract extraction, (c) shows the results of our two step clustering  into 14 classes. (d) shows the results of voxelization  and (e) shows the results of surface extraction. }\label{fig:crop-6}
%\end{figure}
%% if specified like this the section will be ommitted in review mode

>>>>

Frangi et al.~\cite{Frangi1998} introduced a process that searches for geometric structures which are tubular. They define a measure based on two geometric ratios of the second order ellipsoid given by the local Hessian matrix to measure the ``vesselness'' criterion.  In order to determine reliable Hessians, we compute the same metric. We include their work here for completeness and direct the reader to ~\cite{Frangi1998} for details. Let $\lambda_{K}$ be the eigenvalue with the $K^{th}$ smallest magnitude. Here, $|{\lambda}_{1}| \leq| {\lambda}_{2}|\leq| {\lambda}_{3}| $ are the eigenvalues of the Hessian matrix. Specifically, a pixel belonging to a vessel region will have small $\lambda_{1}$ ($|\lambda_{1}|\approx 0$) and $\lambda_{2}$, $\lambda_{3}$ of large magnitude and of equal sign($|\lambda_{1}| \ll |\lambda_{2}|$ and $|\lambda_{2}|\approx |\lambda_{3}|$). The sign indicates if the vessel is bright in a dark background or dark in a bright background. In our case the individual fibers are bright ($\lambda_2,\lambda_3 < 0$). The following measures are defined in~\cite{Frangi1998}.  

@@

The next set of tasks evaluated the spatial context perceived from MetaTracts.\replacedWith{ We achieve this by asking the evaluators}{The participants were asked} to select the fiber bundle which was largest, smallest, or the most similar to a given one from a \removed{sub}set of fiber bundles. Due to the problems of occlusions and similar grayscale values these tests proven to be more difficult to perform successfully with volume rendering of the plain data, compared to MetaTracts. In the Likert scale, the majority strongly agreed that the tasks are undemanding compared to volume rendering.
Finally, the majority of participants agreed that the extracted meshes provides better spatial context than volume rendering.

\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}[tb]
\centering
	\includegraphics[width=\linewidth,  trim = 0mm 0mm 0mm 00mm, clip]{images_pvis/figure9_AMA.pdf}
	 \vspace{-1.5em}
	%\includegraphics[width=0.5\textwidth,  trim = 10mm 40mm 0mm 15mm, clip]{imagesMT2014/image_length_histogram.pdf}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Figure~\ref{fig:orientation_clustering}d, clustered into 10 clusters. The unit for length is the grid cube length.  Clusters labeled 2,3,4,5 and 7 all have cardinality less than 10. }
	\label{fig:len_dist_crop16}
	\vskip-0.2cm
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions, which is either known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$. Large $h$ keeps the major bundles intact. For example, in  Figure~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five clusters, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4,5 and 7 in Figure~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown \added{(avoiding over-segmentation)}.
The following parameters were fixed for all the tests. We set the reliable Hessian threshold $R_{H}$ to be 0.3. A $R_{H}$ of 0.0 would mean all points have reliable local orientation which would cause spurious MetaTracts detection.\replacedWith{On the other hand, a}{A} very high $ R_{H}$ would lead to a decline in number of MetaTracts produced. The $\alpha$ and $\beta$ in $R_{H}$ are as explained in Frangi et al. \cite{Frangi1998} and set to 0.5. The length and the radius parameters for the cylinders of MetaTracts decide how coarse the fiber cylinders are. These are dependent on the underlying fiber characteristic and the weaving pattern. Larger cylinders will handle noisy local orientation better as it inspects a higher number of candidate points to extend the fiber. We used 10.00 and 2.00 for length and radius (measured in grid voxel size), respectively for all tests.  
Simpler geometry (\replacedWith{example dataset 2}{e.g., D2} ) was experimentally found to handle larger cylinders better. $\eta$ in sec.~\ref{subsec:dist_clustering} decides how quickly the hierarchical clustering converges, experimentally values 0.3 to 0.6 removed $1.2\% - 5\%$ of fibers (total number of fibers $\sim$10000) and gave similar results. 
 %Short tracts of less than 0.01 times the smallest axes size are prone to have erroneous local orientation and are therefore removed. 
 $\alpha$ and $\beta$ in equation~\ref{eqn:algo_1} decide how quickly the value of the factor decays; we have used [7-10] and half the length of cylinder (5.0). Our number of fiber bundle directions are limited. Thus even for small $m$ (cardinality of lower dimension in orientation clustering), the distinction between the orientation clusters is preserved quite well. We compared  $m=3-7$ experimentally without any change in results. 
%\begin{figure}[tb]
%	\centering
%	\includegraphics[width=0.5\textwidth, trim = 0mm 80mm 0mm 170mm, clip]{images_pvis/figure10}
%	\caption{Direct hierarchical clustering of MetaTracts Fig.~\ref{fig:length_distribution}(A) by proximity into 10,15 and 20 clusters.}
%	\label{fig:comparison}
%\end{figure}
%

%\begin{figure}
%\centering
% \subfloat[$\eta=7$]{\includegraphics[width=0.3\linewidth]{imagesMetaTracts/Rplot-ndim3k2}}
%  \subfloat[$\eta=3$]{\includegraphics[width=0.3\linewidth]{imagesMetaTracts/Rplot-ndim7k2}}
%  \caption{Embedding from $\mathbb{R}^m\to \mathbb{R}^\eta$, here m is the number of MetaTracts, 580 for this example. Followed by clustering into K=2 clusters.}
%  \label{fig:eta}
%\end{figure}
%\begin{figure}[htp]
%\centering
% \includegraphics[width=\linewidth]{imagesMetaTracts/kul-results.eps}
%  \caption{Results from orientation based dimensionality reduction and clustering. A is the original data set, B shows the results of fiber tracking. C shows the results of orientation based clustering and D shows the results of the proximity based clustering. }\label{fig:kull-full}
%\end{figure}
\subsection{Limitations}
\label{sec:limitations}
\added{
A key assumption of the method is ``connectivity'' (Sec.~\ref{sec:char_data}).
If the ``connectivity'' is not satisfied due to noise in the image, then the  generated MetaTracts will be inaccurate. 
The clustering process also assumes that the fiber bundles have a minimum width.
}
\removed{
In this paper, we introduce a framework to extract and visualize fiber bundles in composite materials. We show that our framework works at comparatively low resolution and with dense fiber arrangements (when extracting single fibers might not be possible). It handles complex fiber patterns such as ``cross overs'' and ``braiding''. We perform an user evaluation to indicate the usefulness of the method.
}

\section{Conclusions and Future Work}
\label{sec:conclusions}
In this paper, we introduce a framework to extract and visualize fiber bundles in composite materials. We show that our framework works at comparatively low resolution and with dense fiber arrangements (when extracting single fibers might not be possible). It handles complex fiber patterns such as ``cross overs'' and ``braiding''.
In future, we plan to increase the precision of the extractions, include uncertainty based visualization to better portray the surface of separations and speed up extraction times.
%
%\begin{figure}
%  \centering
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-scalar.eps}}
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-tracks-1.eps}}
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-tracks-clus-a.eps}}
%    \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-tracks-clus-b.eps}}
%  \subfloat[]{\includegraphics[width=0.2\linewidth]{imagesMetaTracts/crop-6-labels-all.eps}}
%  \caption{Dense data set with bundles which are flat and thin. (a) shows the results original dataset, (b) shows the results of metaTract extraction, (c) shows the results of our two step clustering  into 14 classes. (d) shows the results of voxelization  and (e) shows the results of surface extraction. }\label{fig:crop-6}
%\end{figure}
%% if specified like this the section will be ommitted in review mode
